\documentclass{beamer}
\usetheme{Copenhagen}

\title{Introduction to Python}
\usepackage{listings}
\subtitle{Iterables, Iterators and Generators}
\date{13 November 2015}

\begin{document}
\definecolor{keywords}{RGB}{255,0,90}
\definecolor{comments}{RGB}{0,0,113}
\definecolor{red}{RGB}{160,0,0}
\definecolor{green}{RGB}{0,150,0}
\lstset{language=Python, 
        basicstyle=\ttfamily\small, 
        keywordstyle=\color{keywords},
        commentstyle=\color{comments},
        stringstyle=\color{red},
        showstringspaces=false,
        identifierstyle=\color{green},
        procnamekeys={def,class}}
\begin{frame}
  \titlepage
\end{frame}

\section{Iterables}
\begin{frame}[fragile]{Examples 1}
  \begin{itemize}
	\item {We have already seen that loops in Python are pretty flexible:
	\pause
  \begin{block}{}
\begin{lstlisting}{frame=single}
for i in [1,2,3]:
    print i # prints elements
for	i in "123":
    print i # prints characterwise
\end{lstlisting}
\end{block}
	}
  \end{itemize}
\end{frame}
\begin{frame}[fragile]{Examples 2}
  \begin{itemize}
	\item {We have already seen that loops in Python are pretty flexible (excuse lack of indentation):
  \begin{block}{}
\begin{lstlisting}{frame=single}
for i in xrange(4):
    print i #prints the range without storing
for i in enumerate([1,2,3]):
    print i # gives me indices as well as elements
\end{lstlisting}
\end{block}
	}
	\pause
	\item{ Hopefully you see this and think how do I do that in my code!}
  \end{itemize}
\end{frame}
\begin{frame}[fragile]{Definition}
  \begin{itemize}
	\item { An \textbf{iterable} is any object that responds to the function iter(obj) 
	}
	\pause
	\item{ Python likes to hide this! 
  \begin{block}{}
  \begin{lstlisting}{frame=single}
	for i in x: 
	    # Do loop stuff
  \end{lstlisting}
\end{block}
}
\item Call iter(x) and assign to some name that is never used in the code.
\pause
\item iter(x) returns an object that can do one thing and one thing alone - respond to \_\_next\_\_()
\pause
\item Assign i to the response to \_\_next\_\_() each time round until it throws StopIteration exception (sort of an error) and then move on.
\end{itemize}
\end{frame}
\begin{frame}[fragile]{Definition}
  \begin{itemize}
  \begin{block}{}
  \begin{lstlisting}{frame=single}
	for i in x: 
	    # Do loop stuff
  \end{lstlisting}
\end{block}
	\item{ This is equivalent to:
  \begin{block}{}
  \begin{lstlisting}{frame=single}
  randomName = iter(x)
  while(True):
      try:
          i = randomName.__next__()
          # Do loop stuff
      except StopIteration:
           break
  \end{lstlisting}
\end{block}
}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Uses for Iterables}
  \begin{itemize}
	\item{ Iterables can be used all over the place! 
  \begin{block}{}
  \begin{lstlisting}{frame=single}
  new_list = list(iterable)
  total = sum(iterable)
  smallest = min(iterable)
  largest = max(iterable)
  combined = ''.join(iterable)
  result = [f(x) for x in iterable]
 \end{lstlisting}
\end{block}
}

\end{itemize}
\end{frame}
\section{Generators}
\begin{frame}[fragile]{Make your own iterables}
  \begin{itemize}
  \item{ To make an object and iterable you only need to define the \_\_iter\_\_() hook.}
  \item{\textbf{Generators} are a great way to create your own iterables that are extremely flexible}
  \pause
  \item{You define generators in a similar way to function but using the keyword \textbf{yield}
  \begin{block}{}
\begin{lstlisting}{frame=single}
def myXRange(n):
    i = 0
    while( i < n):
        yield i
        i += 1
\end{lstlisting}
\end{block}
}
\pause
\item This responds to the \emph{next} call by running until it hits yield and then stopping and waiting.
\pause
\item Demo Time!
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Make your own iterables 2}
  \begin{itemize}
\item No reason to make our iterable only finite!
\item{ Lets have all of the positive even numbers
\begin{block}{}
\begin{lstlisting}{frame=single}
def evens():
    i = 2
    while True:
        yield i
        i += 2
\end{lstlisting}
\end{block}
}
\pause
\item Be careful when using these! Don't try to make a list out of this one...
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Make your own iterables 3}
\begin{itemize}
\item{ Perhaps most importantly, we can make iterables that depend on other iterables.
\pause
\begin{block}{}
\begin{lstlisting}{frame=single}
def myEnumerate(iterable):
    index = 0
    for i in iterable:
        yield (index,i)
        index += 1
\end{lstlisting}
\end{block}
}
\item Demo Time!
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Fin}
\begin{itemize}
\item Suggested watching: Loop Like A Native, Ned Batchelder
\item Suggested Library: itertools (This probably has what you need!)
\end{itemize}
\end{frame}



\end{document}



